---
title: "SIRT6_db R Client Walkthrough"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages
To use the SIRT6_db R client, you need to load the required packages first:
```{r load_packages}
library(arrow)
library(dplyr)
```

## Initializing the SIRT6_db Client
Source the client file (adjust the path as necessary):
```{r source_client}
base_path <- "./SIRT6/Aging/PrepareSirt6Metadata/SIRT6_db/"
source(paste0(base_path, "utils/R_client.R"))
```

Initializing the SIRT6_db client:
```{r init_client}
db <- SIRT6db$new(db_path = base_path)
```

## Accessing Database Information
```{r db_info}
cat(paste("Database Version:", db$version(), "\n"))
cat(paste("Available Species:", paste(db$available_species(), collapse = ", "), "\n"))
```

List all available experiments:
```{r list_experiments}
print(db$list_experiments())
```

## Querying Data
You can query the database for specific data. For example, to get a specific experiment:
```{r get_experiment}
data <- db$load_dataset(species = "homo_sapiens", gse_id = "GSE213425")
```

Look what is inside the data object:
```{r data_content}
print(names(data))
```

Accessing the expression data content:
```{r expression_content}
head(data$expression)
```

Accessing the metadata:
```{r samples_content}
head(data$samples)
```

Filter lowly expressed genes using the following criteria: >=5 counts in >=40% samples, <=60% zeros
```{r gene_filtering}
filtered_data <- db$filter_low_expressed_genes(data, min_counts = 5, min_samples_pct = 0.4, max_zeros_pct = 0.6)
```

Create a SummarizedExperiment object for a downstream analysis
```{r create_summarized_experiment}
se <- db$get_summarized_experiment_object(filtered_data)
se
```

## Session Info
```{r session_info}
sessionInfo()
``` 
